---
title: "homework_w1"
author: "Sarah Hvid Andersen"
date: "1/2/2022"
output: html_document
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(echo = TRUE)
```

## Option 1  

```{r}
pacman::p_load(tidyverse, leaflet, htmlwidgets)

```

Loading the data and dropping rows where an NA is present in the latitude or longitude columns. 

```{r}

rf <- read_csv('C:/Users/Sarah/Desktop/R_21/spatial/cds-spatial_fork/Week01/data/RCFeature2018.csv')

rf <- rf %>% drop_na(Latitude)
rf <- rf %>% drop_na(Longitude)

```


Creating the map of blue mountains rainforest and adding the different layers and the minimap.

```{r}
# Set the location and zoom level
leaflet() %>% 
  setView(150.3, -33.75, zoom = 12) %>%
  addTiles()  # checking I am in the right area


# Bring in a choice of esri background layers  

l_bm <- leaflet() %>%   # assign the base location to an object
  setView(150.3, -33.75, zoom = 13)

esri <- grep("^Esri", providers, value = TRUE)

for (provider in esri) {
  l_bm <- l_bm %>% addProviderTiles(provider, group = provider)
}

BMmap <- l_bm %>%
  addLayersControl(baseGroups = names(esri),
                   options = layersControlOptions(collapsed = FALSE)) %>%
  addMiniMap(tiles = esri[[1]], toggleDisplay = TRUE,
             position = "bottomright") %>%
  addMeasure(
    position = "bottomleft",
    primaryLengthUnit = "meters",
    primaryAreaUnit = "sqmeters",
    activeColor = "#3D535D",
    completedColor = "#7D4479") %>% 
  htmlwidgets::onRender("
                        function(el, x) {
                        var myMap = this;
                        myMap.on('baselayerchange',
                        function (e) {
                        myMap.minimap.changeLayer(L.tileLayer.provider(e.name));
                        })
                        }") %>% 
  addControl("", position = "topright")

BMmap

```

Here we are adding the markers. The higher the accuracy was, the larger the circle is in the map. The Esri.WorldImagery background layer allows for a nice visualization of all the markers.

```{r}
BMmap %>% addCircleMarkers(lng = rf$Longitude, 
             lat = rf$Latitude,
             radius = rf$Accuracy,
             popup = paste(
               "<br><strong>FeatureID: </strong>", rf$FeatureID,
               "<br><strong>FeatureType: </strong>",rf$FeatureType,
               "<br><strong>Description: </strong>",rf$Description))

saveWidget(BMmap, "BMmap.html", selfcontained = TRUE)
```


This final version adds clusters to the map. This is not very useful in this case, as the datapoints are so close together. It therefore does not improve the overview of the datapoints which makes it useless. 

```{r}
BMmap %>% addCircleMarkers(lng = rf$Longitude, 
             lat = rf$Latitude,
             radius = rf$Accuracy,
             popup = paste(
               "<br><strong>FeatureID: </strong>", rf$FeatureID,
               "<br><strong>FeatureType: </strong>",rf$FeatureType,
               "<br><strong>Description: </strong>",rf$Description),
             clusterOptions = markerClusterOptions())


```

