---
title: "Working with sf objects"
author: "Sarah Hvid Andersen"
date: "11-February-2022 updated `r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
---

```{r setup, include=FALSE}
pacman::p_load(knitr, rmdformats) 

## Global options
options(max.print="75")
opts_chunk$set(echo=TRUE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

#  Goals: 

1. In the first exercise, you will learn to construct maps with tmaps() by modifying the provided code. The libraries, data, and initial code is provided below.
2. In the second exercise, you learn to: 
 - Understand the provided datasets
 - Learn how to reproject spatial data
 - Limit your data into an area of interest
 - Create a new map
3. In the third exercise, you reproject layers to a shared SRS.
4. In the fourth (and optional) exercise, you create a map of Denmark using online data.

#### Required R libraries

We will use the sf, raster, and tmap packages.
Additionally, we will use the spData and spDataLarge packages that provide new datasets.

```{r}
pacman::p_load(sf, raster, tmap, spData, spDataLarge)

#install.packages("spDataLarge", repos = "https://geocompr.r-universe.dev")

```


#  Exercise 1 - Make a map

In this exercise you will learn to make a map with tmaps library, by adding spatial data layers and modifying the arguments that specify their rendering

#### Data sets  

We will use two data sets: `nz_elev` and `nz`. They are contained by the libraries
The first one is an elevation raster object for the New Zealand area, and the second one is an sf object with polygons representing the 16 regions of New Zealand.

#### Existing code

Here is the code to create a new map of New Zealand (following Robin Lovelace's example).Your role is to improve this map based on the suggestions below.

```{r initial-map}
tm_shape(nz_elev)  +
  tm_raster(title = "elev", 
            style = "cont",
            palette = "BuGn") +
  tm_shape(nz) +
  tm_borders(col = "red", 
             lwd = 3) +
  tm_scale_bar(breaks = c(0, 100, 200),
               text.size = 1) +
  tm_compass(position = c("LEFT", "center"),
             type = "rose", 
             size = 2) +
  tm_credits(text = "A. Sobotkova, 2022") +
  tm_layout(main.title = "My map",
            bg.color = "orange",
            inner.margins = c(0, 0, 0, 0))
```


## Tasks 

1. Change the map title from "My map" to "New Zealand".
2. Update the map credits with your own name and today's date.
3. Change the color palette to "-RdYlGn". 
  (You can also try other palettes from http://colorbrewer2.org/)
4. Put the north arrow in the top right corner of the map.
5. Improve the legend title by adding the used units (m asl).
6. Increase the number of breaks in the scale bar.
7. Change the borders' color of the New Zealand's regions to black. 
  Decrease the line width.
8. Change the background color to any color of your choice.


## Your solution

```{r improved-map}
tm_shape(nz_elev)  +
  tm_raster(title = "Elevation (MASL)", 
            style = "cont",
            palette = "-RdYlGn") +
  tm_shape(nz) +
  tm_borders(col = "black", 
             lwd = 2) +
  tm_scale_bar(breaks = c(0, 50, 100, 150, 200),
               text.size = 1) +
  tm_compass(position = c("RIGHT", "top"),
             type = "rose", 
             size = 2) +
  tm_credits(text = "S. Hvid, 10-02-2022") +
  tm_layout(main.title = "New Zealand",
            bg.color = "lightblue",
            inner.margins = c(0, 0, 0, 0))
```


# Exercise 2 - Inspect and display spatial data

We will use two data sets: `srtm` and `zion` , explore them and make a map.
The first one is an elevation raster object for the Zion National Park area in Israel, and the second one is an sf object with polygons representing borders of the Zion National Park.

Read two new datasets, `srtm` and `zion`, using the code below and then follow the instructions and answer the questions related.

####  Preparation: Load data from provided package
You need to have installed the spDataLarge package for the following code to work. 

```{r load-data}
srtm  <-  raster(system.file("raster/srtm.tif", package = "spDataLarge"))
zion  <-  read_sf(system.file("vector/zion.gpkg", package = "spDataLarge"))
```

## Instructions and questions
Type answers to the questions as code comments next to or under the code used

* Display the `zion` object and view its structure.

  - What can you say about the content of this file?
  - What type of data does it store? 
  - What is the coordinate system used?
  - How many attributes does it contain?
  - What is its geometry?
  
```{r inspect-data-zion}
zion #  it is an sf object with 1 feature and 11 fields. it is a polygon
st_crs(zion) # the coordinate reference system is User input: UTM Zone 12, Northern Hemisphere
crs(zion, asText = TRUE) #"+proj=utm +zone=12 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs"

extent(zion) # xmin: 302903.1 ymin: 4112244 xmax: 334735.5 ymax: 4153087
ncell(zion) #12
plot(zion)
```

* Display the `srtm` object and view its structure.

  - What can you say about the content of this file? 
  - What type of data does it store?
  - What is the coordinate system used? 
  - How many attributes does it contain?
  - How many dimensions does it have? 
  - What is the data resolution?
 
## Your solution and answers

```{r inspect-data}
srtm #class: RasterLayer, dimensions 457, 465, 212505  (nrow, ncol, ncell), res 0.0008333333, 0.0008333333  (x, y)
st_crs(srtm) # WGS 84 (with axis order normalized for visualization) 
crs(srtm, asText = TRUE) #"+proj=longlat +datum=WGS84 +no_defs"
nlayers(srtm) #1 layer
plot(srtm)
```

# Exercise 3 - Reconciling Coordinate systems

Now that you know that coordinate systems differ, make them compatible!


## Instructions

* Reproject the `srtm` dataset into the coordinate reference system used in the `zion` object. 

  - Create a new object `srtm2`
  - Visualize the results using the `plot()` function.

* Reproject the `zion` dataset into the coordinate reference system used in the `srtm` object.
  - Create a new object `zion2`
  - Visualize the results (`zion2` together with `srtm2`) using the `plot()` function.

## Your solution

```{r}
# getting the crs of the Zion object
the_crs <- crs(zion, asText = TRUE)

# creating srtm2, with the crs from zion
srtm2 <- projectRaster(srtm, crs = the_crs, method = 'ngb') 
st_crs(srtm2) #+proj=utm +zone=12 +ellps=GRS80 +units=m +no_defs 

plot(srtm2)


#getting the crs of the srtm object
the_crs2 <- crs(srtm, asText = TRUE)

# creating zion2, with the crs from srtm
zion2 <- st_transform(zion, crs = the_crs2)
st_crs(zion2) #+proj=longlat +datum=WGS84 +no_defs 


# creating plots
plot(srtm2);plot(zion2$geom, add = TRUE) # this does not make sense, they were different to begin with so they still would be now?

#taking the second version and the first version of each data set makes more sense. 
plot(srtm2);plot(zion$geom, lwd = 2, add = TRUE)
plot(srtm);plot(zion2$geom, lwd = 2, add = TRUE)




```


# Exercise 4 - Make a map of Denmark (Optional)

Make a handsome map of Denmark from data available online with a scale, a north arrow and source credits!

## Hints

*Load elevation and regional boundaries for Denmark using the `getData()` function from the `raster` package. 

  - Remember that the `getData()` function takes a number of different attributes, plus the country code in a ISO 3-letter format. You can google what it is for Denmark, or use the `getData('ISO3')` function to view it. For an example of how to use the `getData()` function [see this blog](https://www.gis-blog.com/r-raster-data-acquisition/)
* Make a map with `tmaps` library following the Exercise no.1

## Your solution
```{r}
DKelevation <- getData("alt", country = "DNK", mask = TRUE)
region <- getData("GADM", country = "DNK", level = 1)


tm_shape(DKelevation)  +
  tm_raster(title = "Elevation (MASL)", 
            style = "cont",
            palette = "RdYlGn") +
  tm_shape(region) +
  tm_text("VARNAME_1", size = "AREA", root = 3) +
  tm_borders(col = "black", 
             lwd = 1.5) +
  tm_scale_bar(breaks = c(0, 25, 50, 75, 100),
               text.size = 0.7) +
  tm_compass(position = c("RIGHT", "top"),
             type = "rose", 
             size = 3) +
  tm_credits(text = "S. Hvid, 11-02-2022") +
  tm_layout(main.title = "Denmark",
            bg.color = "lightblue",
            inner.margins = c(0, 0, 0, 0),
            legend.text.size = 0.55)
```

